# Data Sources:
- Spotify Data: Kaggle Spotify Countries dataset link: https://www.kaggle.com/datasets/asaniczka/top-spotify-songs-in-73-countries-daily-updated/universal_top_spotify_songs.csv
- GDP Data: IMF Website

# DS 202 Final Project : Spotify Data by Country Analysis Using R

Authors: Logan Jorgensen, David Szczepanik, Devdatta Kasture

# Introduction

The goal of this report is to analyze Spotify's listening data by country/region, and see if there are any interesting trends or correlations between the data and other sources. For our particular report, we are comparing the Spotify listening data against the GDP of a country, and investigating if there are any noticeable strong trends between a country's GDP and the music its people listen to. Although the data we have does not include genre/subgenres, Spotify has a few metrics it applies to each song in their database which can give us an understanding of what kind of music is being listened to in a specific country. We shall explore some of these metrics within our project below.

# Data
## Importing Libraries Used in our Report

Down below are the libraries that we used within our investigation. We used ggplot2 for data visualization, readxl for reading in our excel sheet, dplyr for data manipulation and countrycode, which was used due to a discrepancy in two-character and three-character country codes.

```{r, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)

install.packages("readxl", repos="https://cran.r-project.org")
library(readxl)
library(dplyr)

install.packages("countrycode", repos="https://cran.r-project.org")
library(countrycode)
```


## Importing the IMF GDP-by-Country Data
The GDP country data from the IMF website is made available in an excel workbook (.xls) format. The following code extracted the data from the sheet in the file called "Data", and stored it in a dataframe called imf_data for our investigation. We made some adjustments to only include the data that we needed, such as skipping the first three rows of data that were not useful for our investigation, as can be seen below.

```{r}

# Read in the Excel file with the read_excel() function. The data desired is in the "Data" sheet. The first three rows are info for the user, so we skip them.
#  Column names are in the fourth row, so we set the col_names argument to TRUE.

imf_data <- read_excel("imf_data.xls", sheet = "Data", skip = 3, col_names = TRUE)

# View the first 6 rows of the data

```

## Filtering the IMF GDP-by-Country Data from the Excel Workbook
Once we had extracted the data from the excel workbook, we had a useable dataframe to work off of. Our next step would be to prepare this dataframe by cleaning the data and pre-processing the dataframe for a merge with the spotify data in the future. We only used data from the most recent year avaiable to study only the most recent information available, which in this case was for the year 2021. We extracted the data for each country for that respective year, and renamed the column in our dataframe to 2021 GDP per capita for descriptiveness.

```{r}
# The only columns we need are the `Country Code` and `2021` columns.
filtered_imf_gdp_data <- imf_data[, c("Country Code", "2021")]

# Next, we can rename the columns to be more descriptive (rename "2021" to "2021 GDP per capita")
colnames(filtered_imf_gdp_data) <- c("Country Code", "2021 GDP per capita")

# Display the first 6 rows of the filtered data
head(filtered_imf_gdp_data)
```

## Combining Extra Metadeta from the IMF Excel Workbook
Within the same GDP Excel workbook from earlier, there was another sheet named "Metadata - Countries", which contained exactly what it is named - the metadata for each country. It included the region, income level, any special notes, and the full name for every country, which not only provides more context to some of these countries, but could be also used for further exploratory analysis. We took the metadata and stored it in it's own dataframe, before merging it with the dataframe above to create one large dataframe of exactly the information we needed to move on: Country Code, GDP per capita (2021), Region, Income Group, Special Notes, and Table Name. We shall move on to the spotify data next.

```{r}
# Read in the Excel file, this time from the sheet "Metadata - Countries"
imf_metadata <- read_excel("imf_data.xls", sheet = "Metadata - Countries")

# We want to merge the rows of the "Metadata - Countries" into the "filtered_imf_data" dataframe based on the "Country Code" column.
# We can use the left_join() function from the dplyr package to do this.

merged_imf_data <- left_join(filtered_imf_gdp_data, imf_metadata, by = "Country Code")

# Display the first 6 rows of the merged IMF data
head(merged_imf_data)
```


## Pulling in the Spotify Data
Now that the GDP data has been gathered, filtered, and made ready to join the spotify data, we move on to doing the same thing for the spotify data. The spotify dataset has been downloaded from Kaggle, and is stored under the name "universal_top_spotify_songs.csv". We started off by reading this data into a dataframe.

```{r}
# Read in the Spotify data
spotify_data <- read.csv("universal_top_spotify_songs.csv")

# Display the first 6 rows of the Spotify data
head(spotify_data)
```


## Dealing With Inconsistent Country Code Data
This is where we came across an issue that required the countrycode library. There was a discrepancy between the two datasets: the IMF dataset uses Alpha-3 ISO codes for countries, while the Spotify dataset uses Alpha-2 ISO codes. Below we convert both the Alpha-3 codes, and the Alpha-2 codes, all to their country names using the library. This ensured we have the country names associated to the data going forward in our exploration, rather than country codes.

```{r}

# Convert the Alpha-2 codes in the Spotify data `country` column to Alpha-3 codes to match the IMF data in a new dataframe copy of the Spotify data
spotify_data_alpha3 <- spotify_data %>%
  mutate(country = countrycode(country, "iso2c", "iso3c"))

```


## Joining the IMF and Spotify Data
Now that we've converted the country codes in the Spotify (when possible) to match the IMF data, we should be able to join the two datasets.


### Checking for Overlapping Country Codes
We checked to see if there were any matches between the country codes between the two datasets. As you can see below, we have valid workable data between 71 countries, more than enough to explore possible coorelations.

```{r}

# Show the country codes that overlap between spotify_data_alpha3's `country` column and merged_imf_data's `Country Code` column
intersect(spotify_data_alpha3$country, merged_imf_data$`Country Code`)

```

### Joining the Two Datasets on the Country Code
Based on the output from above, we were able to join the two datasets on the `Country Code` from the GDP dataframe and `country` from the Spotify dataframe, to create one big dataframe for our project. This concludes our data cleaning, and move on to exploring the merged data and beginning to study the results.

```{r}
# Code to join the two datasets on the `Country Code` and `country` columns, and store the result in a new dataframe called `joined_data`.
joined_data <- left_join(spotify_data_alpha3, merged_imf_data, by = c("country" = "Country Code"))

# Remove any data where the `country` column is <NA>
cleaned_joined_data <- joined_data[!is.na(joined_data$country), ]

# Show the first 6 rows of the joined data
head(cleaned_joined_data)
```

<!-- Another sample / test just to explore the joined data -->
```{r}
# Display the first 6 rows of the joined data for `country`, `2021 GDP per capita` and `artists` columns to show that the join merged the data correctly
head(cleaned_joined_data[, c("country", "2021 GDP per capita", "artists")])
```




<!-- End of Data Cleaning and Preprocessing -->
---------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
<!-- Beginning of Data Exploration and Analysis -->

# Data Exploration and Analysis
Below is an example of how we can extract and analyze data from our new dataframe; We can not only quickly create a dataframe to hold the most popular song in the country with the code ZAF, but display it and chart it all within the same group of code. This is how we will explore the data further into our report.
  
```{r}
# For the data in each row with `country` equal to "ZAF" with the `daily rank` column equal to 1, show the `name` and `snapshot_date` column
zaf_most_popular_song <- cleaned_joined_data %>%
  filter(country == "ZAF" & daily_rank == 1) %>%
  select(name, snapshot_date)

# Display the first 6 rows of the data
head(zaf_most_popular_song)

# Create a visual that counts how many times each song was the most popular song in ZAF
ggplot(zaf_most_popular_song, aes(x = name)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Song Name", y = "Number of Days as Most Popular Song", title = "Most Popular Song in ZAF by Day")
```

## Number of Explicit vs. Non-Explicit Song Appearances by Country
Our first investigation was to compare how many explicit vs. non-explicit songs each country listens to. We wanted to explore if there would be a correlation between a country's GDP and explicitness of the music a country listens to. The first visual below, as you can probably tell, was very hard to look at, so we ended up creating a second visual which included the same data, but rather than displaying both the number of explicit and non-explicit songs, we ended up only displaying the percentage of explicit songs over the count of total songs for each country. We also colored in the bar for the United States for visual purposes.

```{r, fig.width=20, fig.height=10}
# We can create a new dataframe that counts the number of explicit and non-explicit songs for each country (true or false in is_explicit column). Exlcude entries with country value of <NA>.
explicit_vs_non_explicit <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(country, is_explicit) %>%
  summarise(count = n())

# Display the first 6 rows of the data
head(explicit_vs_non_explicit)

# Then, we create a visual that shows the number of explicit and non-explicit songs for each country
ggplot(explicit_vs_non_explicit, aes(x = country, y = count, fill = is_explicit)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Country", y = "Number of Songs", title = "Number of Explicit vs. Non-Explicit Songs by Country")

```

```{r, fig.width=20, fig.height=10}
# We can add a new column that represents the percentage of songs that are explicit for each country
explicit_percent_by_country <- explicit_vs_non_explicit %>%
  group_by(country) %>%
  mutate(percent = count / sum(count) * 100)

# Then, we create a visual that shows the percentage column value, where the y axis is the percentage of songs that are explicit for each country and the x axis is the country.
# The column that has the country value of "USA" should have it's bar colored red.

ggplot(explicit_percent_by_country, aes(x = country, y = percent, fill = country)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Country", y = "Percentage of Songs that are Explicit", title = "Percentage of Explicit Songs by Country") +
  scale_fill_manual(values = c("USA" = "darkred"))
```

## Tempo by Country
After exploring the explicitness data per country, we moved onto the average tempo of the top music per country. Generally speaking, songs with a higher tempo are often times more upbeat and share positive characteristics, while slower tempo music is often more relaxed and share neutral/negative characteristics. This is an excellent metric to study trends in music per country. Again, as was the case above, the chart was very hard to read, so we categorized the same tempo data by the region of each country and averaged them out over each region instead. 

```{r, fig.width=20, fig.height=10}
# We can create a new dataframe that calculates the average tempo for each country with it's region value. Excluding entries with country value of <NA>.
tempo_by_country <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(country, Region) %>%
  summarise(avg_tempo = mean(tempo))

# Then, we create a visual that shows the average tempo for each country
# Color the bar for the country based on it's `Region` value using fill
ggplot(tempo_by_country, aes(x = country, y = avg_tempo, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Country", y = "Average Tempo", title = "Average Tempo by Country")
```

## Tempo by Region

```{r, fig.width=20, fig.height=10}
# We can create a new dataframe that calculates the average tempo for each region. Exlcuding entries with country value of <NA>.
tempo_by_region <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(Region) %>%
  summarise(avg_tempo = mean(tempo))

# Then, we create a visual that shows the average tempo for each region, colored by region
ggplot(tempo_by_region, aes(x = Region, y = avg_tempo, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Region", y = "Average Tempo", title = "Average Tempo by Region")
```

## Danceability by GDP per Capita
We can also explore one of Spotify's custom song metrics for coorrelation between GDP and listening trends. Danceability, according to Spotify's API, describes how suitable a track is for dancing based on a combination of musical elements. We kept Danceability on our report due to a bunch of repetition in results, but we also explored energy, speechiness, and acousticness to extremely similar results, and decided to omit for conciseness. Below is a chart that compares the Danceability metric against GDP.

```{r, fig.width=20, fig.height=10}
# We need get the average of danceability and the average "2021 GDP per capita" for each country. Excluding entries with country value of <NA>. Then we will plot the data as a scatterplot, with the x-axis being the average "2021 GDP per capita" and the y-axis being the average danceability.
danceability_and_gdp_by_country <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(country) %>%
  summarise(avg_danceability = mean(danceability), avg_gdp = mean(`2021 GDP per capita`))

# Then, we create a visual that shows the average danceability for each country, colored by region
ggplot(danceability_and_gdp_by_country, aes(x = avg_gdp, y = avg_danceability, color = country)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "2021 GDP per capita", y = "Average Danceability", title = "Average Danceability by GDP per Capita")
```

## Danceability by GDP per Capity Quartile
Building off of our Danceability investigation above, we furthered our exploration by dividing the GDP into quartiles, and created two charts based off of the quartile split rather than GDP value itself. Although the scatter plot is identical to the one above, the box plot which was not available to us before dividing up the GDP into quartiles gives us an insight to some coorrelation that no one expected. More on this in Results.

```{r, fig.width=20, fig.height=10}
danceability_and_gdp_by_country <- cleaned_joined_data %>%
  filter(!is.na(country) & !is.na(`2021 GDP per capita`)) %>%
  group_by(country) %>%
  summarise(avg_danceability = mean(danceability, na.rm = TRUE), 
            avg_gdp = mean(`2021 GDP per capita`, na.rm = TRUE))

# Calculate quartiles for GDP per capita
gdp_quartiles <- quantile(danceability_and_gdp_by_country$avg_gdp, probs = c(0.25, 0.5, 0.75), na.rm = TRUE)

# Add a new column for GDP quartiles
danceability_and_gdp_by_country$gdp_quartile <- cut(danceability_and_gdp_by_country$avg_gdp, 
                                                    breaks = c(-Inf, gdp_quartiles, Inf), 
                                                    labels = c("1st Quartile", "2nd Quartile", "3rd Quartile", "4th Quartile"),
                                                    include.lowest = TRUE)

# Scatterplot colored by GDP quartile
ggplot(danceability_and_gdp_by_country, aes(x = avg_gdp, y = avg_danceability, color = gdp_quartile)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "2021 GDP per capita", y = "Average Danceability", title = "Average Danceability by GDP per Capita Quartile")

```

```{r}
# Create box plots for each GDP quartile
ggplot(danceability_and_gdp_by_country, aes(x = gdp_quartile, y = avg_danceability)) +
  geom_boxplot() +
  labs(x = "GDP per Capita Quartile", y = "Average Danceability", title = "Danceability by GDP per Capita Quartile") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Valance by GDP per Capita
We decided to plot Valance over GDP as well to include a metric that had a visible difference from Danceability. Valance, according to the Spotify API, is a measure from 0 to 100 describing the musical positiveness conveyed by a track. Again, it's a metric developed by Spotify, but it can be explored just as all the other ones.

```{r, fig.width=20, fig.height=10}
# Now, we do the same thing as above, but instead of danceability, we use valence
valence_and_gdp_by_country <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(country) %>%
  summarise(avg_valence = mean(valence), avg_gdp = mean(`2021 GDP per capita`))

# Then, we create a visual that shows the average valence for each country, colored by region
ggplot(valence_and_gdp_by_country, aes(x = avg_gdp, y = avg_valence, color = country)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "2021 GDP per capita", y = "Average Valence", title = "Average Valence by GDP per Capita")
```

# Results

After exploring the data using our methods above, we found some incredible things that no one in our group expected. 

Starting off from the Percentage of Explicit Songs by Country bar chart, we were originally caught off-guard by how low the percentage of explicit music was for the United States compared to the rest of the world. It clicked when we took into consideration the volume of music that is created in the United States, both explicit and non-explicit, that helped justify the lower than average percentage. We were also slightly blown away that India had approximately 100% of it's listening trends labelled as explicit. 

Looking at Tempo by Country didn't give us exactly concrete information to work off of, but Logan's idea of splitting up the data from country-based to region-based gave us one key outlier, and something to keep in mind going forward with the rest of the regions. The key outlier, South Asia, averaged a tempo of roughly 110 Beats per Minute, way lower than the 120 Beats per Minute that every other region hovered around. This could be used further to investigate whether there is a cultural influence in that region which leads to listening trends of slower music, such as genre preference or enjoyment of music.

The greatest shock from this project can be located in the Average Danceability by GDP per Capita plot. As a team, we were not really expecting a strong coorrelation between music and GDP, but we were all taken by surprise when we saw the strong correlation between the danceability of a song and GDP. We can confidently present that there is a strong coorrelation between a higher danceability and lower GDP, which is further established by the boxplot we created when splitting the data into quartiles. The first and second quartiles were above a score of 0.7, while third was at 0.7 and fourth was closer to 0.6 than 0.7. The same trends were followed by energy, speechiness, and acousticness metrics.

The last chart we created compared Valance to GDP. We included this chart to show that some music metrics do not seem to be influenced by GDP. It can be noted that higher GDP has lower average valence but even at lower GDP, the results are so scattered we can confidently say there is no coorrelation between Valance (musical positiveness) and GDP.

# Conclusion

After exploring the data and concluding that there are some variables in music that are influenced by GDP, this leaves us in an excellent position moving forward. As stated earlier, some variables such as danceability, energy, etc. are influenced by GDP, which can be used to further our exploration. However, some things need to be addressed before furthering our studies; Countries with low GDP tend to be dominated by one artist, and some variables are not influenced by GDP. Some studies could be done to examine the music industries in lower GDP countries, and more data never hurts to have regarding the Spotify metrics that we examined earlier. Perhaps as more data becomes available, some data may begin to smooth out coorrelations or create new ones. If we were to continue working on this investigation, we would begin by simplifying the data into region-based analysis, and perhaps examining how Spotify recommends new music to its users or how music is being created in countries with differing economic stature. Overall, we were more than pleased to discover there is some coorrelation, and more than proud of the work we had done to reach our conclusion.
