# DS 202 Final Project Proposal
## Analysis of Spotify Data by Country Using R

### Project Overview
The objective of this project is to explore and analyze Spotify data across different countries, focusing on identifying trends or correlations between music preferences and other socio-economic factors such as a country's Gross Domestic Product (GDP). This analysis aims to uncover insights into how economic conditions may influence musical tastes and preferences.

### Data Sources
1. **Spotify Data**: The dataset is obtained from Kaggle, titled "Top Spotify Songs in 73 Countries - Daily Updated". This dataset includes daily rankings of songs across various countries.
   - Source: [Kaggle Spotify Countries Dataset](https://www.kaggle.com/datasets/asaniczka/top-spotify-songs-in-73-countries-daily-updated/universal_top_spotify_songs.csv)
2. **GDP Data**: The International Monetary Fund (IMF) provides GDP data for various countries, available on their website. This data will be used to analyze correlations with Spotify's music data.

### Initial Data Cleaning

#### Importing and Cleaning the IMF GDP Data

```{r}
library(readxl)
imf_data <- read_excel("imf_data.xls", sheet = "Data", skip = 3, col_names = TRUE)
filtered_imf_gdp_data <- imf_data[, c("Country Code", "2021")]
colnames(filtered_imf_gdp_data) <- c("Country Code", "2021 GDP per capita")
head(filtered_imf_gdp_data)
```

#### Importing and Preprocessing the Spotify Data

```{r}
spotify_data <- read.csv("universal_top_spotify_songs.csv")
head(spotify_data)
```


#### Dealing With Inconsistent Country Code Data
There is a discrepency between the two datasets: the IMF dataset uses Alpha-3 ISO codes for countries, while the Spotify dataset uses Alpha-2 ISO codes. We will need to convert the Alpha-3 codes and Alpha-2 codes to the actual country names, and then join the two datasets.

```{r}

# Convert the Alpha-2 codes in the Spotify data `country` column to Alpha-3 codes to match the IMF data in a new dataframe copy of the Spotify data
spotify_data_alpha3 <- spotify_data %>%
  mutate(country = countrycode(country, "iso2c", "iso3c"))

# Display the first 6 unique values in the `country` column of the Spotify data that aren't <NA>
unique(spotify_data_alpha3$country, incomparables = FALSE)
```


#### Joining the IMF and Spotify Data
Now that we've converted the country codes in the Spotify (when possible) to match the IMF data, we should be able to join the two datasets.


##### Checking for Overlapping Country Codes
First, it's probably a good idea to check that the country codes in the two datasets match up. We can do this by checking for any country codes that are in both datasets:
```{r}

# Show the country codes that overlap between spotify_data_alpha3's `country` column and merged_imf_data's `Country Code` column
intersect(spotify_data_alpha3$country, merged_imf_data$`Country Code`)

```

##### Joining the Two Datasets on the Country Code
Based on this output, we should be able to join the two datasets on the `Country Code` in  merged_imf_data and `country` in spotify_data_alpha3.

```{r}
# Code to join the two datasets on the `Country Code` and `country` columns, and store the result in a new dataframe called `joined_data`.
joined_data <- left_join(spotify_data_alpha3, merged_imf_data, by = c("country" = "Country Code"))

# Remove any data where the `country` column is <NA>
cleaned_joined_data <- joined_data[!is.na(joined_data$country), ]

# Show the first 6 rows of the joined data
# head(cleaned_joined_data)

# Display the first 6 rows of the joined data for `country`, `2021 GDP per capita` and `artists` columns to show that the join merged the data correctly
head(cleaned_joined_data[, c("country", "2021 GDP per capita", "artists")])
```

### Marginal Summaries

#### Summary of the Spotify Data
## Number of Explicit vs. Non-Explicit Song Appearances by Country

```{r, fig.width=20, fig.height=10}
# We can create a new dataframe that counts the number of explicit and non-explicit songs for each country (true or false in is_explicit column). Exlcude entries with country value of <NA>.
explicit_vs_non_explicit <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(country, is_explicit) %>%
  summarise(count = n())

# Display the first 6 rows of the data
head(explicit_vs_non_explicit)

# Then, we create a visual that shows the number of explicit and non-explicit songs for each country
ggplot(explicit_vs_non_explicit, aes(x = country, y = count, fill = is_explicit)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Country", y = "Number of Songs", title = "Number of Explicit vs. Non-Explicit Songs by Country")

```

This visual is a bit hard to look at, so instead we can create a visual that shows the percentage of explicit songs as a percentage of all songs for each country.

```{r, fig.width=20, fig.height=10}
# We can add a new column that represents the percentage of songs that are explicit for each country
explicit_percent_by_country <- explicit_vs_non_explicit %>%
  group_by(country) %>%
  mutate(percent = count / sum(count) * 100)

# Then, we create a visual that shows the percentage column value, where the y axis is the percentage of songs that are explicit for each country and the x axis is the country.
# The column that has the country value of "USA" should have it's bar colored red.

ggplot(explicit_percent_by_country, aes(x = country, y = percent, fill = country)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Country", y = "Percentage of Songs that are Explicit", title = "Percentage of Explicit Songs by Country") +
  scale_fill_manual(values = c("USA" = "darkred"))
```

## Tempo by Country
We can also look at the average tempo of songs by country.

```{r, fig.width=20, fig.height=10}
# We can create a new dataframe that calculates the average tempo for each country with it's region value. Excluding entries with country value of <NA>.
tempo_by_country <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(country, Region) %>%
  summarise(avg_tempo = mean(tempo))

# Then, we create a visual that shows the average tempo for each country
# Color the bar for the country based on it's `Region` value using fill
ggplot(tempo_by_country, aes(x = country, y = avg_tempo, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Country", y = "Average Tempo", title = "Average Tempo by Country")
```

## Tempo by Region
We can also look at the average tempo of songs by region overall, instead of breaking it down by country.

```{r, fig.width=20, fig.height=10}
# We can create a new dataframe that calculates the average tempo for each region. Exlcuding entries with country value of <NA>.
tempo_by_region <- cleaned_joined_data %>%
  filter(!is.na(country)) %>%
  group_by(Region) %>%
  summarise(avg_tempo = mean(tempo))

# Then, we create a visual that shows the average tempo for each region, colored by region
ggplot(tempo_by_region, aes(x = Region, y = avg_tempo, fill = Region)) +
  geom_bar(stat = "identity", position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Region", y = "Average Tempo", title = "Average Tempo by Region")
```

